---
title: "Example 4.1"
author: "Kevin Cummiskey"
date: "October 15, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Example 4.1 Recovering Polyphenols (pg 272)

## Lesson 16 - Simple Linear Regression

Objectives:

1. Describe association between two quantivative variables.

2. Interpret least-squares regression models.

3. Compare and contrast separate versus linear regression models.

What is the observational unit? explanatory variable? response variable?

\vspace{1in}

What type of study is this?

\vspace{1in}

```{r}
grapes = read.table(file = "http://www.isi-stats.com/isi2/data/Polyphenols.txt", header = T)
grapes %>% ggplot(aes(x = Ethanol., y = PC)) + geom_point()
```

Write a statistical model for a separate mean for each ethanol level.

<br><br><br>

Fit the model.

```{r}
grapes$Ethanol_cat = factor(grapes$Ethanol.)
contrasts(grapes$Ethanol_cat) = contr.sum
model_anova = lm(PC ~ Ethanol_cat, data = grapes)
summary(model_anova)
anova(model_anova)
```

What would you conclude from this model?

\vspace{1in}

Instead, let's say we fit the following regression model:

\[y_{i} = \beta_0 + \beta_1 x_i + \epsilon_{i}\]
\[\epsilon_{ij} \sim \text{N}(0,\sigma^2)\]
where $i = 1,\ldots,15$ is grape, $x_i$ is the ethanol concentration used on the $i$th grape, and $y_i$ is the PC of the $i$th grape.

How is this model different that the ANOVA model?

\vspace{1in}

Let's fit a regression model.  

```{r}
model_regression = lm(PC ~ Ethanol., data = grapes)
summary(model_regression)
anova(model_regression)
```

What would you conclude from this output? What has changed from the ANOVA model?

\vspace{1in}

## Lesson 17 - Inference for Simple Linear Regression

Objectives:

1. Simulation-based inference for relationship between quantiative variables. 

2. Theory-based approach for relationship between quantiative variables.

3. Evaluate validity conditions for theory-based tests.

Why do we want to conduct inference?

\vspace{1in}


Write the null and alternative hypothesis to test whether there is an association between ethanol concentration used and PC.

\vspace{1in}

### Simulation-based approach

Describe how we would conduct a simulation to conduct this test.

\vspace{1in}


Let's conduct a simulation-based test (pg 291).

```{r}
m = 5000 #number of iterations
slopes = c() # empty vector
grapes.sim = grapes #copy of data

for(i in 1:m){
  grapes.sim$PC.sim = sample(grapes.sim$PC) #shuffle the response
  model.sim = lm(PC.sim ~ Ethanol., data = grapes.sim) # fit model to shuffled data
  slopes[i] = coef(model.sim)[2] # extract the slope from the model
}
```


Here is a plot of the distribution of the simulated slopes:

```{r}
hist(slopes)
```

Here is the $p$-value:
```{r}
sum(slopes > coef(model_regression)[2])/m
```

What would we conclude?

\vspace{1in}

### Theory-based test

The $p$-value is in the linear model object:
```{r}
summary(model_regression)
```

A $t$ confidence interval for the population slope $\beta_1$ is (pg295):

\vspace{1in}

```{r}
b1 = summary(model_regression)$coefficients[2,1]
se_b1 = summary(model_regression)$coefficients[2,2]
tstar = qt(0.975,13)

upper = b1 + tstar * se_b1
lower = b1 - tstar * se_b1

lower 
upper

# or you can just do
confint(model_regression)
```






