---
title: "Section 5.2 Observational Studies with Multiple Quantitative Variables"
author: "Kevin Cummiskey"
date: "November 4, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Review

Let's say you are interested in the effects of sleep and coffee consumption on academic performance.  You randomly assign cadets in a class to one of three sleep groups (3,5,7 hours) and coffee groups (0, 200, 400 mg) in a 3 x 3 factorial design and record their performance on a WPR.  Let $y_i$ be the WPR grade, $x_{1,i}$ be sleep in hours (standardized), and $x_{2,i}$ be the coffee consumption (standardized) of cadet $i$ . Consider the following model:

\[y_i = \beta_0 + \beta_1 x_{1,i} + \beta_2 x_{2,i} + \epsilon_i \quad \epsilon_i \sim N(0, \sigma^2)\]

\begin{enumerate}
\item Why do we standardize explanatory variables?
\item Interpret $\beta_0$.
\item Will the adjusted and unadjusted effect of sleep on WPR performance be the same? Explain.
\item Can we use this model to test whether the effect of sleep on WPR performance depends upon coffee consumption? Explain.
\end{enumerate}

# Example 5.2

```{r, message=FALSE}
library(tidyverse)
brains = read.table(file = "http://www.isi-stats.com/isi2/data/BrainSize.txt",
                    header = T)
```

## PIQ and Brain Size

```{r, fig.height=3}
brains %>% ggplot(aes(x = BrainSize, y = PIQ)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
model_unadjusted = lm(PIQ ~ BrainSize, data = brains)
summary(model_unadjusted)
```

What conclusions should we draw from the unadjusted model?

\vspace{2in}

## Confounding by Height?

```{r,fig.height=3, message=F}
library(gridExtra)
p1 = brains %>% ggplot(aes(x = BrainSize, y = Height)) + 
  geom_point() + geom_smooth(method = "lm", se = F)
p2 = brains %>% ggplot(aes(x = Height, y = PIQ)) + 
  geom_point() + geom_smooth(method = "lm", se = F)
grid.arrange(p1,p2, ncol = 2)
```

Is there evidence of confounding? How will adjusting for height change the effect of BrainSize on PIQ?

\vspace{1in}

Let's look at the height-adjusted PIQ.

```{r, fig.height = 3, warning = FALSE}
model_PIQheight = lm(PIQ ~ Height, data = brains)
summary_PIQheight = model_PIQheight %>% fortify()

#height-adjusted PIQ
brains = brains %>% 
  mutate(PIQ_height_adj = mean(PIQ) + summary_PIQheight$.resid)

#plots on page 382
# color scale
sc = scale_color_distiller(palette = "Spectral", direction = 1) 

p3 = brains %>% ggplot(aes(x = BrainSize, y = PIQ, color = Height)) +
  geom_point() + ylim(65,155) + sc
p4 = brains %>% ggplot(aes(x = BrainSize, y = PIQ_height_adj, color = Height)) +
  geom_point() + ylim(65,155) + sc
grid.arrange(p3,p4,ncol = 2)
```

How did we obtain height-adjusted PIQ? What happened to PIQ after adjusting for height?

\vspace{1in}

Let's look at height-adjusted brainsize.

```{r, fig.height = 3}
model_Brainheight = lm(BrainSize ~ Height, data = brains)
summary_Brainheight = model_Brainheight %>% fortify()

#height-adjusted PIQ
brains = brains %>% 
  mutate(BrainSize_height_adj = mean(BrainSize) + summary_Brainheight$.resid)

#plots on page 382
p5 = brains %>% ggplot(aes(x = BrainSize, y = PIQ, color = Height)) +
  geom_point() + ylim(65,155) + sc
p6 = brains %>% 
  ggplot(aes(x = BrainSize_height_adj, y = PIQ_height_adj, color = Height)) +
  geom_point() + 
  ylim(65,155) +  
  sc
grid.arrange(p5,p6,ncol = 2)
```

How do we obtain height-adjusted BrainSize? 

\vspace{1in}

Let's look at an added variable plot predicting PIQ from BrainSize before (black line) and after (blue line) adjusting for height.

```{r, fig.height=3}
p6   + geom_smooth(method = "lm", se = FALSE) +
  geom_abline(aes(slope = 0.1177, intercept = 4.66))
```

Let's fit the main effects model.

```{r}
model_height_brainsize = lm(PIQ ~ BrainSize + Height, data = brains)
summary(model_height_brainsize)
```

What would we conclude from these results?

\vspace{1in}

## Interaction

If we test for an interaction, what research question are we answering?

\vspace{0.25in}

Why should we standardize BrainSize and Height?

```{r}
brains = brains %>% mutate(std.BrainSize = scale(BrainSize),
                           std.Height = scale(Height))
model_interaction = lm(PIQ ~ std.BrainSize * std.Height,
                       data = brains)
summary(model_interaction)
```

What do we conclude from these results?

\vspace{1in}


