---
title: "Example4_3"
author: "Kevin Cummiskey"
date: "October 18, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


Read in the data and perform data analysis.

```{r, fig.height = 3}
houses = read.table(file = "http://www.isi-stats.com/isi2/data/housing.txt",
                    header = TRUE)
head(houses)
houses %>% ggplot(aes(x = sqft, y = price.1000, color = lake)) + geom_point() +
  geom_smooth(method = "lm")
```

What are the observational units? explanatory variable(s)? outcome variable?

\vspace{1in}

What is the study design?

\vspace{0.25in}

Let's fit a simple linear regression model.

\[y_i = \beta_0 + \beta_1 x_i + \epsilon_i \qquad \epsilon_i \sim N(0,\sigma^2)\]

where $y_i$ is the price and $x_i$ is the size (sq ft) of house $i$. 

```{r, warning=FALSE, message = FALSE, fig.height=3}
model_simple = lm(price.1000 ~ sqft, data = houses)
summary(model_simple)
```

What is the interpretation of $\beta_1$?

\vspace{1in}

Let's look at the residuals?

```{r, fig.height = 3}
houses = houses %>% mutate(residuals_simple = residuals(model_simple))
houses %>% ggplot(aes(x = residuals_simple, fill = lake)) + geom_histogram()
houses %>% ggplot(aes(x = sqft, y = residuals_simple, color = lake)) + geom_point()
houses %>% ggplot(aes(x = lake, y = residuals_simple, color = lake)) + geom_point()
``` 

What do these residuals tell us?

\vspace{1in}

Next, let's consider a model including location.

\[y_i = \beta_0 + \beta_1 x_{1,i} + \beta_2 x_{2,i} + \epsilon_i \qquad \epsilon_i \sim N(0,\sigma^2)\]

where $y_i$ is the price of house $i$, $x_{1,i}$ is the size (sq ft) of house $i$, and $x_{2,i}$ is 1 if house $i$ is lakefront and is 0 otherwise. 

$x_{2,i}$ is an indicator variable.  How else could we encode this variable?

\vspace{0.5in}

What assumptions does this model make?

\vspace{2in}

Let's fit the model.

```{r}
#Reverse coding of lake
houses$lake = factor(houses$lake, levels = c("notlakefront",
                                             "lakefront"))

model_withLake = lm(price.1000 ~ sqft + lake, data = houses)
summary(model_withLake)
anova(model_withLake)
```

What is the interpretation of $\beta_1$?

\vspace{1in}

What conclusions would you make from this model?

\vspace{1in}

Let's look at the residuals.

```{r, warning=FALSE}
houses = houses %>% mutate(residuals_lake = residuals(model_withLake))
houses = houses %>% mutate(fitted_lake = fitted(model_withLake))
houses %>% ggplot(aes(x = fitted_lake, y = residuals_lake, color = lake)) + 
  geom_point()
```

What do the residuals tell us?

\vspace{1in}



