---
title: "ACFT"
author: "Kevin Cummiskey"
date: "September 23, 2019"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Read in the data and perform data analysis.


```{r cars}
acft = read_csv("https://raw.githubusercontent.com/nick3703/MA376/master/ACFT.csv")

acft %>% ggplot(aes(x = Surface, y = Sled)) + geom_boxplot()

```


# Analysis 1 - Fit a model for surface effect

\[y_{ij} = \mu + \alpha_j + \epsilon_{ij}\]

where $\alpha_j$ is the effect of surface $j$.

This is a bad thing to do as our model assumes iid error, which is not the case because we have multiple measurements on each subject.

```{r}
acft$Surface = factor(acft$Surface, levels = c("G","S","T"))
contrasts(acft$Surface) = contr.sum
single.lm = lm(Sled ~ Surface, data = acft)
summary(single.lm)
anova(single.lm)
```

Note the large $p$-value for Surface 1.


# Simulation based approach for Analysis 1

```{r}
f.sim = c()
m = 5000

for(i in 1:m){
  acft$Surface.sim = sample(acft$Surface)
  sim.model = lm(Sled ~ Surface.sim, data = acft)
  f.sim[i] = anova(sim.model)$"F value"[1]
}

hist(f.sim)
```


Now, let's look at person effects.

```{r}
summary = acft %>% group_by(Participant) %>% summarise(mean = mean(Sled))
summary %>% ggplot(aes(x = Participant, y = mean)) + geom_point()
```


Let's fit a model that includes person effects

\[y_{ij} = \mu + \alpha_j + \beta_i + \epsilon_{ij}\]

where $\alpha_j$ is the effect of surface $j$ and $\beta_i$ is the effect of person $i$ (also called the block effect).

```{r}
acft$Participant = factor(acft$Participant)
contrasts(acft$Participant) = contr.sum

#IMPORTANT - don't use factor in Partipant below
# Doing so will reset the contrasts you set in contr.sum

adj.mod = lm(Sled ~ Surface + Participant, data = acft)
summary(adj.mod)
anova(adj.mod)

```

Note the smaller $p$-values after adjusting for block effects.

# Analysis 2 - revising the simulation to adjust for blocking

Now, we only shuffle individuals time:

```{r}
f.adj = c()
acft.mod = acft

for(i in 1:m){
  acft.mod = acft.mod %>% group_by(Participant) %>% sample_n(3)
  acft.mod$shuffled.cat = rep(c("S","T","G"), 25)
  shuff.lm = lm(Sled ~ shuffled.cat, data = acft.mod)
  f.adj[i] = anova(shuff.lm)$"F value"[1]
}

hist(f.adj)
hist(f.sim)


```


Analysis 3 - adjust each response for the block effect

```{r}
#calculate the block effect
acft = acft %>% group_by() %>% mutate(mean_overall = mean(Sled))
acft = acft %>% group_by(Participant) %>% mutate(mean_block = mean(Sled))
acft = acft %>% group_by(Participant) %>% mutate(block_effect = mean_block - mean_overall)

#calculate the adjusted sled time
acft = acft %>% group_by() %>% mutate(Sled.adj = Sled - block_effect)
```

Compare the adjusted sled times to the original sled times.  Note they have the same means but different standard deviations.  We have adjusted for the between person variability.

```{r}
#Compare adjusted sled time to sled time
acft %>% ggplot(aes(x = Surface, y = Sled.adj)) + geom_boxplot()

acft %>% group_by(Surface) %>% summarise(mean = mean(Sled), sd = sd(Sled))
acft %>% group_by(Surface) %>% summarise(mean = mean(Sled.adj), sd = sd(Sled.adj))
```

Fit a model for adjusted sled time with surface as a predictor.

```{r}
adjusted.lm = lm(Sled.adj ~ Surface, data = acft)
summary(adjusted.lm)
anova(adjusted.lm)
anova(single.lm)
```






