---
title: "FloJo"
author: "Kevin Cummiskey"
date: "9/29/2019"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Read in the data and perform basic data analysis

```{r}
cred = read.table(file = "http://www.isi-stats.com/isi2/data/CorporateCredibility.txt",
                  header = TRUE)

cred %>% group_by(Treatments) %>% summarise(mean(PI), sd(PI), n())
cred %>% ggplot(aes(x = Treatments, y = PI)) + geom_boxplot()
```

One-variable analysis (p202) of purchase intent by treatment (this is not the best idea).

1. You cannot answer the question in terms of the explanatory variables

2. You unnecessarily lose degrees of freedom.

```{r}
cred$Endorser = factor(cred$Endorser)
cred$CorpCred = factor(cred$CorpCred)
contrasts(cred$Endorser) = contr.sum
contrasts(cred$CorpCred) = contr.sum

oneway.lm = lm(PI ~ Treatments, data = cred)
summary(oneway.lm)
anova(oneway.lm)

```

Fit a two-way ANOVA. Note the SS for CorpCred and SS for Endorser are the same as they would have been if we fit a one-way ANOVA for each of these variables.  This occurs because of the balanced design.

```{r}
twoway.lm = lm(PI ~ CorpCred + Endorser, data = cred)
summary(twoway.lm)
anova(twoway.lm)
```

Let's look at the residuals

```{r}
cred$residuals.twoway = residuals(twoway.lm)
cred$fitted.twoway = fitted.values(twoway.lm)

cred %>% ggplot(aes(x = residuals.twoway)) + geom_histogram()
cred %>% ggplot(aes(x = fitted.twoway, y = residuals.twoway)) + geom_point()

```



