---
title: "Example4_4"
author: "Kevin Cummiskey"
date: "October 24, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Review

In Chapter 4, we've been looking at housing prices in Michigan. 

```{r, fig.height = 3}
houses = read.table(file = "http://www.isi-stats.com/isi2/data/housing.txt",
                    header = TRUE)
houses %>% 
  ggplot(aes(x = sqft, y = price.1000, color = lake)) +
  geom_point()
```

In Example 4.3, we looked at the effect of house size on price after adjusting for location.  Here is the model:

\[price_i = \alpha_0 + \alpha_1 sqft_i + \alpha_2 lake_i + \epsilon_i \qquad \epsilon_i \sim N(0,\sigma^2)\]

where $price_i$ is the price of house $i$, $sqft_i$ is the size (sq ft) of house $i$, and $lake_i$ is 1 if house $i$ is lakefront and is 0 otherwise. 

What assumptions does this model make?

\vspace{1in}

How do we interpret $\alpha_1$? $\alpha_2$?

\vspace{1in}

```{r, fig.height = 3}
#Reverse coding of lake so 1 is lake and 0 is not lake
houses$lake = factor(houses$lake, levels = c("notlakefront",
                                             "lakefront"))

model_withLake = lm(price.1000 ~ sqft + lake, data = houses)

houses %>% 
  mutate(price.predict = predict(model_withLake, newdata = .),
         residuals = price.1000 - price.predict) -> houses

houses %>% 
  ggplot(aes(x = sqft, y = price.predict, color = lake)) +
  geom_line() +
  geom_point(aes(y = price.1000))

summary(model_withLake)
```

Report estimates of model parameters here.

\vspace{0.75in}

Calculate the expected price of a 2500 sq ft house that is not on the lake front.

\vspace{0.75in}

Calculate the expected price of a 2500 sq ft house on the lake front.

\vspace{0.75in}

Let's take a look at the residuals vs the predicted (fitted) values.

```{r, fig.height=3}
houses %>% ggplot(aes(x = price.1000, 
                      y = residuals,
                      color = lake)) +
  geom_point() + geom_hline(yintercept = 0)
```

Do we see evidence of an interaction here?

\vspace{1in}

## Two models

If we suspect the price per square foot differs for lake front and nonlake front, we could fit two separate simple models.

```{r, fig.height=3}
houses %>% ggplot(aes(x = sqft, y = price.1000, color = lake)) +
  geom_point() + geom_smooth(method = "lm")

lake_model <- lm(price.1000 ~ sqft, data = houses %>% filter(lake == "lakefront"))
coef(lake_model)

nonlake_model <- lm(price.1000 ~ sqft, data = houses %>% filter(lake == "notlakefront"))
coef(nonlake_model)
```

## Interactions

Let's look at a model with an interaction.

\[price_i = \beta_0 + \beta_1 sqft_i + \beta_2 lake_i + \beta_3 sqft_ilake_i + \epsilon_i \qquad \epsilon_i \sim N(0,\sigma^2)\]

How do we interpret $\beta_1$? $\beta_2$? $\beta_3$? $\beta_1 + \beta_3$?

\vspace{1.5in}

```{r, fig.height=3}
model_interaction = lm(price.1000 ~ sqft * lake, data = houses)
summary(model_interaction)
anova(model_interaction)
```

Write the equation for the price of a lake front house as a function of size.

\vspace{0.5in}

Write the equation for the price of a nonlake house as a function of size.

\vspace{0.5in}

How do these equations compare to the two models above? Why is the interaction model preferable?

\vspace{0.5in}

Is there evidence the price per square foot is different for lakefront and nonlakefront homes? How do you know?

\vspace{1in}

What would you conclude from these results? Your answer should include discussion of effect sizes, significance, and overall predictive capability of the model.

\vspace{2in}

The main effect of location is not significant in the model with an interaction.  Should we conclude the location of the house is not associated with price? Justify your answer.

\vspace{1in}

```{r, fig.height=3}
anova(model_interaction)
```

Note: The ANOVA table from R will not match the textbook because there are different ways to calculate sums of squares when the data is unbalanced (as in observational studies). The textbook reports Type III ANOVA tables in this article \url{https://mcfromnz.wordpress.com/2011/03/02/anova-type-iiiiii-ss-explained/}



For Type III ANOVA, you can use the \tt{car} package in R. Note the capital ``A" in the function below.

```{r, warning=FALSE, message=FALSE}
library(car)
Anova(model_interaction, type = 3)
```
