---
title: "Intro to Logistic Regression"
author: "Kevin Cummiskey"
date: "November 18, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
library(tidyverse)
```

# Section 6.2 Introduction to Logistic Regression

## Review

In the article ``The incidence of thyroid disorders in the community: A twenty-year follow-up of the Wickham survey" by Vanderpump et al., 443 of the 582 smokers and 502 of the 732 nonsmokers were still alive at the 20 year follow-up.  

Calculate the odds ratio and logs odds ratio of being alive comparing smokers to nonsmokers.

\vspace{0.5in}

Perform the chi-square test to determine if there is a significant association between smoking and survival.  State the appropriate hypotheses and report the chi-square test statistic and $p$-value.

\vspace{1in}

Does your result above mean smoking raises the probability of being alive?

\vspace{0.5in}

## Useful log rules for this lesson

- $e^{\ln r} = r$
- $\ln (e^r) = r$
- $e^r \times e^s = e^{r+s}$
- $\ln\left(\frac{r}{s}\right) = \ln(r)-\ln(s)$


## Introduction to Logistic Regression

```{r, fig.height = 3}
smoke = read.table(file = "smoke.csv", header = T, sep = ",")
smoke$Alive = factor(smoke$Alive)
smoke$Smoker = factor(smoke$Smoker)
smoke %>% group_by(Smoker,Alive) %>% count() %>% spread(key = "Smoker",value = "n")
smoke %>% ggplot(aes(x = Age, y = Smoker, color = Alive)) + geom_jitter(width = 2, height = 0.2) 
```

Based on the plot, is age a confounding variable of the association between smoking and survival? Explain.

\vspace{1in}

### Let's look at the relationship between survival and age.

```{r, fig.height = 3}
#Let's look at a plot of age vs proportion alive
summary = smoke %>% count(Age, Alive) %>%
  group_by(Age) %>%
  mutate(Alive_proportion = n/sum(n),
         total = sum(n)) %>%
  filter(Alive == 1) %>% select(-c(Alive,n))

summary %>%
  ggplot(aes(x = Age, y = Alive_proportion)) +
  geom_point()
```

Explain two reasons a linear model for age and alive proportion is not appropriate?

\vspace{1in}

Let's try a log odds (or \textit{logit}) transformation.

```{r, fig.height=3}
summary = summary %>% 
  mutate(logOdds = log(Alive_proportion/(1-Alive_proportion)))
summary %>% ggplot(aes(x = Age, y = logOdds)) +
  geom_point()
```

Is a linear model on the log odds appropriate? How does the logit transformation ensure probabilities are between 0 and 1?

\vspace{1in}

Let's fit the following model.

\[\log\left(\frac{\pi_i}{1-\pi_i}\right) = \beta_0 + \beta_1 \text{Age}_i \]

How do we interpret $\beta_0$ and $\beta_1$? Would we expect $\beta_1$ to be positive or negative? Why is there no $\epsilon_i$ on this model?

\vspace{1in}

```{r}
# In practice, we would just do this:
model_age = glm(Alive ~ Age, data = smoke, family = "binomial")
summary(model_age)
```

What is the predicted odds ratio associated with a one-year increase in age?

\vspace{0.25in}

What is the predicted odds ratio associated with a ten-year increase in age?

\vspace{0.25in}

\textbf{Key idea: the slope of a logistic regression model indicates a multiplicative change in the odds.}

What would we conclude from these results?  

\vspace{1in}

How do we find predicted probabilities of survival for each subject?

\vspace{1in}

```{r}
summary_model = model_age %>% fortify()
summary_model = summary_model %>% mutate(predicted_prob = exp(.fitted)/(1 + exp(.fitted)))
```

### Let's look at the relationship between smoking and survival

```{r}
model_smoker = glm(Alive ~ Smoker, data = smoke, family = "binomial")
summary(model_smoker)
```

How do we interpret $\beta_0$ and $\beta_1$? Have we seen these values before?

\vspace{1in}
