---
title: "Example6_3"
author: "Kevin Cummiskey"
date: "November 25, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

# Section 6.3 Multiple Logistic Regression

## Review 

In the article ``The incidence of thyroid disorders in the community: A twenty-year follow-up of the Wickham survey" by Vanderpump et al., 443 of the 582 smokers and 502 of the 732 nonsmokers were still alive at the 20 year follow-up.

Let $Y_i$ be 1 if subject $i$ survives 20 years and 0 if subject $i$ does not survive such that $Y_i \sim \text{Bernoulli}(\pi_i)$. Consider the following model:

\[\log\left(\frac{\pi_i}{1-\pi_i}\right) = \beta_0 + \beta_1 x_i\]

where $x_i$ equals 1 if subject $i$ is a smoker and 0 if subject $i$ is not a smoker.

Interpret the following quantities and provide estimates based on the data.

* $\beta_0$

* $\exp(\beta_0)$

* $\frac{\exp(\beta_0)}{1 + \exp(\beta_0)}$

* $\beta_1$
* $\exp(\beta_1)$
* $\beta_0 + \beta_1$
* $\exp(\beta_0 + \beta_1)$
* $\frac{\exp(\beta_0 + \beta_1)}{1 + \exp(\beta_0 + \beta_1)}$

## Multiple Logistic Regression

Why is it important to adjust for age in this analysis?

\vspace{0.25in}

Here is a plot of the unadjusted association between smoking and 20-year survival.

```{r, fig.height = 3}
smoke = read.table(file = "smoke.csv", header = T, sep = ",")
smoke$Alive = factor(smoke$Alive)
smoke$Smoker = factor(smoke$Smoker)

summary = smoke %>% group_by(Smoker, Alive) %>% 
  count() %>%  group_by(Smoker) %>% mutate(prop = n/sum(n))

summary %>% ggplot(aes(x = Smoker, y = prop, fill = Alive)) +
  geom_col()
```

Let's fit the logistic regression model in the review. Confirm the model produces the same odds ratio as the sample statistics.

```{r}
model.crude = glm(Alive ~ Smoker, data = smoke, family = "binomial")
summary(model.crude)
```

Let's look at the survival of smokers versus nonsmokers with the same levels of age.

```{r}
summary2 = smoke %>% group_by(Age,Smoker, Alive) %>% 
  count() %>%  group_by(Age,Smoker) %>% mutate(prop.alive = n/sum(n)) %>%
  filter(Alive == 1) %>% select(Age,Smoker,prop.alive) 

summary2 %>% ggplot(aes(x = Age, y = prop.alive, color = Smoker)) +
  geom_point()
```

Does it appear smokers or nonsmokers have a better odds of surviving for 20 years?

\vspace{0.5in}

Let's fit the following model where $Y_i \sim \text{Bernoulli}(\pi_i)$.

\[\log\left(\frac{\pi_i}{1-\pi_i}\right) = \beta_0 + \beta_1 x_{1,i} + \beta_2 x_{2,i}\]

where $x_{1,i}$ is smoker (1) or nonsmoker (0) and $x_{2,i}$ is the subject's age at the start of the study.

```{r}
#remove older folks - no survivors 75 years or older.
smoke_cleaned = smoke %>% filter(Age < 70)

model.age_adjusted = glm(Alive ~ Smoker + Age, data = smoke_cleaned, family = "binomial")
summary(model.age_adjusted)

library(car)
Anova(model.age_adjusted, type = "III")
```

How does the adjusted odds ratio compare to the unadjusted?

\vspace{1in}


