---
title: "Example 1_2practice"
author: "Kevin Cummiskey"
date: "September 4, 2019"
output: rmarkdown::github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
odor = read.table(file = "OdorRatings.txt",
                  header = T)
```

Let's look at the data:

```{r}
odor %>% ggplot(aes(x = condition, y = rating)) +
  geom_boxplot()
```

Let's conduct some inference

In words, we would want to test these hypotheses:

$H_0$: There is no association between being exposed to the scent and the rating of the store.

$H_a$: There is an association between being exposed to the scent and the rating of the store.

First, let's do a randomization test

```{r}
#observed statistic
odor %>% group_by(condition) %>% summarise(mean(rating), sd(rating))

odor %>% group_by(condition) %>% summarise(mean(rating)) %>%
  summarise(diff(`mean(rating)`))
```

Question: if there was no effect of scent, then how unusual would a difference of 1.3 in ratings be?

We can assess this with a randomization test.

```{r}
#number of iterations to perform
m = 10000
sim.diff = c()
sim.t = c()
for(i in 1:m){
  sim.dat = odor %>% mutate(sim.condition = sample(condition))
  sim.scent =sim.dat %>% filter(sim.condition == "scent")
  sim.noscent = sim.dat %>% filter(sim.condition == "noscent")
  sim.diff[i] = mean(sim.scent$rating) - mean(sim.noscent$rating)
  
  sim.dat = sim.dat %>% group_by(sim.condition) %>% mutate(mean2 = mean(rating))
  sim.dat = sim.dat %>% group_by() %>% mutate(residual = rating - mean2)
  
  sim.SEresiduals = sum(sim.dat$residual^2)/46
  
  sim.t[i] = sim.diff[i]/(sim.SEresiduals*sqrt(1/24 + 1/24))
}

hist(sim.diff, breaks = 100)

hist(sim.t, breaks = 100)

sum(sim.diff > 1.3)/m

```



