---
title: "Omega3"
author: "Kevin Cummiskey"
date: "September 10, 2019"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
diet = read.table(file = "http://www.isi-stats.com/isi2/data/FishOmega3.txt",
                  sep = "\t",
                  header = TRUE)

```

Let's perform some data analysis.

```{r}
#descriptive statistics
diet %>% group_by(fish) %>% summarise(mean(omega.3),
                                      sd(omega.3),
                                      n())

diet %>% ggplot(aes(y = omega.3)) + geom_boxplot()
diet = diet %>% filter(omega.3 < 0.08)

diet %>% ggplot(aes(x = fish, y = omega.3)) + geom_boxplot()
```


Fit a single mean model

\[y_{ij} = \mu + \epsilon_{ij}\]

```{r}
diet = diet %>% mutate(mean1 = mean(omega.3))
SST = sum((diet$omega.3-diet$mean1)^2)
```

Fit a separate means model

\[y_{ij} = \mu_j + \epsilon_{ij}\]

```{r}
diet = diet %>% group_by(fish) %>% mutate(mean2 = mean(omega.3))

#SSE
SSE = sum((diet$omega.3 - diet$mean2)^2)

#SSR
SSR = SST - SSE

#R^2
R2 = SSR/SST

#Fstat
Fstat = (SSR/4)/(SSE/22)
#Question: when is the Fstat big?


model = lm(omega.3 ~ fish, data = diet)
anova(model)
summary(model)

```

Inference on $R^2$. Simulation-based approach.

$H_0: \mu_{A} = \cdots = \mu_{E}$
$H_a: \text{at least one of the means not equal to the others}$


```{r}
m = 5000
rsquared = c() #vector for simulated R2
fstat.sim = c() #vector for simulated Fstat

#Simulation
for(i in 1:m){
  diet$sim.fish = sample(diet$fish) #shuffle fish categories
  diet = diet %>% group_by(sim.fish) %>% mutate(mean2.sim = mean(omega.3)) #calculate new means
  SSE.sim = sum((diet$omega.3 - diet$mean2.sim)^2) #calculate new SSE
  rsquared[i] = 1-SSE.sim/SST #calculate new R2; Question: why don't we calculate SST everytime
  fstat.sim[i] = ((SST-SSE.sim)/4)/(SSE.sim/22) #calculate Fstat
}

#Calculate p-value for R2 = 0.51
hist(rsquared, breaks = 50)
sum(rsquared > R2)/m

#Calculate p-value for Fstat = 5.62
hist(fstat.sim)
x = seq(0,10,0.1)
plot(x, df(x,4,22), type = "l")

1 - pf(5.62, 4, 22)
summary(model)
```


Confidence and Prediction Intervals for the Group Means

```{r}
summary = diet %>%group_by(fish) %>% 
  summarise(mean =mean(omega.3),
            sd = sd(omega.3),
            n = n(),
            residuals2 = sum((omega.3-mean2)^2)
                                                 )
summary = summary %>% mutate(pooledSD = sqrt(sum(residuals2)/(27-5)))

#calculate confidence interval
summary = summary %>% mutate(LBmean = mean - qt(0.975,26)*pooledSD/sqrt(n))
summary = summary %>% mutate(UBmean = mean + qt(0.975,26)*pooledSD/sqrt(n))

#calculate prediction interval
summary = summary %>% mutate(LBprediction = mean - qt(0.975,26)*pooledSD*sqrt(1+1/n))
summary = summary %>% mutate(UBprediction = mean + qt(0.975,26)*pooledSD*sqrt(1+1/n))

summary

```

Consider a one-sample t-test

$H_0: \mu = 8$

$H_a: \mu \neq 8$

\[t = \frac{\bar{x} - \mu_0}{s/\sqrt{n}}\]

```{r}
power.t.test(n = 18, delta = 1, sd = 2, power = NULL, type = "one.sample", alternative = "two.sided")

power.t.test(n = NULL, delta = 1, sd = 2, power = 0.80, type = "one.sample", alternative = "two.sided")

```


Now, let's consider our case:

Cohen's F statistic

\[f = \sqrt{\frac{R^2}{1-R^2}}\]

```{r}
library(pwr)

R2 = 0.3
#Cohen's f statistic
f = sqrt(R2/(1-R2))

pwr.anova.test(k = NULL, n = 10, f = f, power = 0.80)
```








