---
title: "Example4_5"
author: "Kevin Cummiskey"
date: "October 28, 2019"
output: pdf_document
---

# Review

Let $y_i$ and $x_{1,i}$ be quantative variables and $x_{2,i}$ be a categorical variable with two levels.

Model 1: $y_i = \beta_0 + \beta_1x_{1,i} + \beta_2 x_{2,i} + \epsilon_i$

Model 1: $y_i = \alpha_0 + \alpha_1x_{1,i} + \alpha_2 x_{2,i} + \alpha_3 x_{1,i}x_{2,i} + \epsilon_i$

For each of the following tests, which model and parameter would you use?

\begin{enumerate}
  \item There is a linear association between $x_{2}$ and y after adjusting for $x_1$.
  \item There is a linear association between $x_1$ and $y$ for subjects in the reference group of $x_2$.
  \item There is a linear assocation between $x_1$ and $y$ for subjects not in the reference group of $x_2$.
  \item The effect of $x_1$ on $y$ differs by level of $x_2$.
\end{enumerate}

Why are we typically not interested in $\alpha_2$? 

\vspace{0.5in}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# One variable analyses

```{r}
diamonds = read.table(file = "http://www.isi-stats.com/isi2/data/diamonds.txt",
                      header = T)
```



## Price vs Weight

```{r, fig.height=3}
diamonds %>% ggplot(aes(x = Carat, y = Price..1000s.)) +
  geom_point() + geom_smooth(method = "lm", se = F)
model_weight = lm(Price..1000s. ~ Carat, data = diamonds)
summary(model_weight)
anova(model_weight)
```

What would you conclude from this model?

\vspace{1in}


## Price vs clarity

```{r, fig.height = 3}
diamonds %>% ggplot(aes(x = Clarity, y = Price..1000s.)) +
  geom_point()
```

How many indicator variables do we need?

\vspace{0.25in}

Here is the model:

\[y_{i} = \beta_0 + \beta_1 VS1_i + \beta_2 VS2_i + \beta_3 VVS1 + \beta_4 VVS2 + 
\epsilon_i \quad \epsilon_i \sim N(0,\sigma^2)\]

What is the reference category?

\vspace{0.25in}


```{r}
model_Clarity = lm(Price..1000s. ~ Clarity, data = diamonds)
summary(model_Clarity)
anova(model_Clarity)
```

What would you conclude from this model? Why are all the clarity coefficients negative?

\vspace{1in}

# Two variable analysis

## Are there differences in price across clarity categories after adjusting for weight?

Here is the model:

\[y_{i} = \beta_0 + \beta_1 Carat_i + \beta_2 VS1_i + \beta_3 VS2_i + \beta_4 VVS1 + \beta_5 VVS2 + 
\epsilon_i \quad \epsilon_i \sim N(0,\sigma^2)\]

```{r}
model_ClarityWeight = lm(Price..1000s. ~ Carat + Clarity, data = diamonds)
summary(model_ClarityWeight)
anova(model_ClarityWeight)
```

How have the coefficients changed from the one-variable models?

\vspace{1in}

Write out the regression equations for the five categories of clarity. What is the relationship among these regression lines?

\vspace{1in}

```{r, fig.height=3, warning=F}
diamonds = diamonds %>% 
  mutate(predicted2 = predict(model_ClarityWeight, diamonds))
diamonds %>% ggplot(aes(x = Carat, y = Price..1000s., color = Clarity)) +
  geom_point() + geom_line(aes(y = predicted2))
```

Draw inference.

$H_0:$ There is no linear association between clarity and price, after adjusting for diamond weight.

$H_a:$ There is a linear association between clarity and price, after adjusting for diamond weight.

Why not use the $p$-values for each indicator variable?

\vspace{0.5in}

Perform the partial $F$-test (pg 344). 

\vspace{2in}

What do the $p$-values of the indicator variable coefficients tell us?

\vspace{0.5in}

## Does the price increase associated diamond weight differ across clarity categories?

$H_0:$ There is no interaction between clarity and price, after adjusting for diamond weight and diamond clarity.

$H_a:$ There is an interaction between clarity and price, after adjusting for diamond weight and diamond clarity.

```{r, fig.height=3}
model_interaction = lm(Price..1000s. ~ Carat*Clarity, data = diamonds)
summary(model_interaction)
anova(model_interaction)
diamonds %>% ggplot(aes(x = Carat,y = Price..1000s., color=Clarity)) +
  geom_point() + geom_smooth(method = "lm", se = F, fullrange = T)
```

Perform the partial $F$-test.

\vspace{2in}

Confidence Intervals

```{r}
confint(model_interaction)
```



