---
title: "Chapter 4.5 - Multilevel Categorical Variables"
author: "Kevin Cummiskey"
date: "October 28, 2019"
output: pdf_document
---

# Review

Let $y_i$ and $x_{1,i}$ be quantitative variables and $x_{2,i}$ be a categorical variable with two levels.

Model 1: $y_i = \beta_0 + \beta_1x_{1,i} + \beta_2 x_{2,i} + \epsilon_i$

Model 2: $y_i = \alpha_0 + \alpha_1x_{1,i} + \alpha_2 x_{2,i} + \alpha_3 x_{1,i}x_{2,i} + \epsilon_i$

For each of the following tests, which model and parameter would you use?

\begin{enumerate}
  \item There is a linear association between $x_{2}$ and y after adjusting for $x_1$.
  \item There is a linear association between $x_1$ and $y$ for subjects in the reference group of $x_2$.
  \item There is a linear association between $x_1$ and $y$ for subjects not in the reference group of $x_2$.
  \item The effect of $x_1$ on $y$ differs by level of $x_2$.
\end{enumerate}

Let's say you fit Model 2 and you find $\alpha_2$ is not sign-ficant.  Should you conclude there is no association between $x_{2}$ and $y$? Explain.

\vspace{1in}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Diamonds

When buying a diamond, you focus on the 4 C's: cut, clarity, color, and carat (weight).  In this example, we will investigate the association between price, weight, and clarity.  There are five clarity categories (IV, VVS1, VVS2, VS1, VS2).

```{r}
diamonds = read.table(file = "http://www.isi-stats.com/isi2/data/diamonds.txt",
                      header = T)
```

# One variable analyses

## Price vs Weight

Let $y_i$ be the price and $x_i$ be the weight (carats) of diamond $i$.  Write a linear model for price as a function of weight.

\vspace{0.5in}

```{r, fig.height=3}
diamonds %>% ggplot(aes(x = Carat, y = Price..1000s.)) +
  geom_point() + geom_smooth(method = "lm", se = F)
model_weight = lm(Price..1000s. ~ Carat, data = diamonds)
```

Describe two interesting features of the plot above.

\vspace{0.5in}

```{r}
summary(model_weight)
anova(model_weight)
```

How accurately does this model typically predict diamond price?

\vspace{0.25in}

Based on this model, what's a reasonable price per carat?

\vspace{0.25in}

Explain how confounding by clarity could occur in this scenario.

\vspace{0.75in}

## Price vs clarity

We can express models for clarity and price in different ways.  The models below are equivalent.

Write a \underline{separate means} model for price and clarity. \vspace{0.25in}

Write a model for price and clarity using \underline{effect encoding}. \vspace{0.25in}

Write a model for price and clarity using \underline{indicator encoding}. \vspace{0.25in}

```{r, fig.height = 3}
#reorder the levels from highest to lowest clarity
diamonds <- diamonds %>% 
  mutate(Clarity = factor(Clarity, 
                          levels = c("IF", "VVS1", "VVS2", "VS1", "VS2"))) 
#summary statistics
diamonds %>% 
  group_by(Clarity) %>% 
  summarize(mean = mean(Price..1000s.),
            sd = sd(Price..1000s.),
            n = n())
#plot
diamonds %>% ggplot(aes(x = Clarity, y = Price..1000s.)) +
  geom_point()
```

Let's move forward using indicator encoding. Let's fit the model.

```{r}
model_Clarity = lm(Price..1000s. ~ Clarity, data = diamonds)
summary(model_Clarity)
anova(model_Clarity)
```

What is the predicted price for an IF diamond?

\vspace{0.25in}

How accurately does this model typically predict diamond price?

\vspace{0.25in}

Using the results above, find estimates of the parameters in the effect encoding model.

\vspace{1in}

### Weight vs. Carat

The last thing we might want to look at before proceeding to the two-variable analysis is the relationship between clarity and weight.  For time sakes, we'll just look at the means.

```{r}
diamonds %>% 
  group_by(Clarity) %>% 
  summarize(mean = mean(Carat),
            sd = sd(Carat),
            n = n())
```

# Two variable analysis

In this section, we want to:

* Estimate the price per carat after adjusting for clarity.

* Determine whether the clarity is associated with price after adjusting for diamond weight.

Using indicator encoding, write a model for price as a function of weight and clarity.

\vspace{1in}

```{r}
model_ClarityWeight = lm(Price..1000s. ~ Carat + Clarity, data = diamonds)
summary(model_ClarityWeight)
anova(model_ClarityWeight)
```

Based on this model, what's a reasonable price per carat? Has it changed much after adjusting for clarity?

\vspace{1in}

How much does clarity improve your predictions of price?

\vspace{0.5in}

Write out the regression equations for the five categories of clarity. What is the relationship among these regression lines?

\vspace{2in}

```{r, fig.height=3, warning=F}
diamonds = diamonds %>% 
  mutate(predicted2 = predict(model_ClarityWeight, diamonds))
diamonds %>% ggplot(aes(x = Carat, y = Price..1000s., color = Clarity)) +
  geom_point() + geom_line(aes(y = predicted2))
```

Draw inference.

$H_0:$ There is no linear association between clarity and price, after adjusting for diamond weight.

$H_a:$ There is a linear association between clarity and price, after adjusting for diamond weight.

Why not use the $p$-values for each indicator variable?

\vspace{0.5in}

Perform the partial $F$-test (pg 344). 

\vspace{2in}

What do the $p$-values of the indicator variable coefficients tell us?

\vspace{0.5in}

# Interaction

Write a research question we could answer by including the interaction between clarity and weight.

\vspace{2in}


$H_0:$ There is no interaction between clarity and weight.

$H_a:$ There is an interaction between clarity and weight.

```{r, fig.height=3}
model_interaction = lm(Price..1000s. ~ Carat*Clarity, data = diamonds)
summary(model_interaction)
anova(model_interaction)
diamonds %>% ggplot(aes(x = Carat,y = Price..1000s., color=Clarity)) +
  geom_point() + geom_smooth(method = "lm", se = F, fullrange = T)
```

Perform the partial $F$-test.

\vspace{2in}

Confidence Intervals

```{r}
confint(model_interaction)
```



