---
title: "Modeling Nonlinear Associations"
author: "Kevin Cummiskey"
date: "November 6, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Review

Explain a situation in which you would use the partial $F$-test.  As part of your answer, describe your research question, the variables you would collect, the models you would fit, and hypotheses you would test.

\vspace{1.5in}

# Example 5.3 Arctic Sea Ice

The purpose of this analysis is to predict when there will be no sea ice in Artic Ocean.

```{r, fig.height = 3, message = F, warning= F}
library(gridExtra)
ice = read.table(file = "http://www.isi-stats.com/isi2/data/ArticSeaIce.txt",
                 header = T)
model_linear = lm(SASIE_12 ~ Year12, data = ice)

p1 = ice %>% ggplot(aes(x = Year12, y = SASIE_12)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = F, fullrange = T) + 
  ylim(0,8) + 
  xlim(1978, 2020)
p2 = model_linear %>% fortify() %>%
  ggplot(aes(x = .fitted, y = .resid)) + geom_point() +
  geom_hline(yintercept = 0)
grid.arrange(p1,p2, ncol = 2)
```

Is a linear model appropriate? What are the implications of choosing the wrong model?

\vspace{0.75in}


Let's fit the following model:

\[y_i = \beta_0 + \beta_1 x_i + \beta_2 x_i^2 + \epsilon_i \quad \epsilon_i \sim N(0,\sigma^2)\]

How do we interpret the coefficients?
\vspace{1in}

```{r}
# Let's try to fit the model
model_quadratic = lm(SASIE_12 ~ Year12 + I(Year12^2), data = ice)
summary(model_quadratic)
```

What's unusual in the output?

\vspace{1in}


```{r}
# Let's standardize year first
ice = ice %>% mutate(std.year = (Year12-mean(Year12))/sd(Year12))
model_std.quadratic = lm(SASIE_12 ~ std.year + I(std.year^2), data = ice)
summary(model_std.quadratic)
```

Why do we need to standardize year first?

\vspace{1in}

```{r, message = F, warning = F, fig.height=3}
ice = ice %>% left_join(model_std.quadratic %>% fortify())
ice %>% ggplot(aes(x = Year12, y = SASIE_12)) + 
  geom_point() + geom_smooth(method = "lm", se = F) +
  geom_line(aes(y = .fitted), col = "red") + 
  ylim(0,8) + 
  xlim(1978,2020)
```


Let's look at the second-order term.

```{r}
anova(model_std.quadratic, model_linear)
```

What is the name of the test above?

\vspace{0.25in}

What hypotheses are being tested? 

\vspace{0.25in}

What do you conclude from this test?

\vspace{1in}


What year does each model predict there will be no more sea ice in the Arctic Ocean?
\vspace{1in}


